---
title: "Visualizations and Performance Metrics"
output: html_document
date: "2024-12-13"
---

# Project Overview

This project involves building a predictive model for estimating the **Sold Price** of properties based on two characteristics: **Year Built** and **Total Square Footage**. I trained a **k-Nearest Neighbors (kNN)** model, evaluated its performance, and visualized its results.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(randomForest)
library(caret)
library(tidyverse)


FM_housing <- read_csv(
  file = "https://raw.githubusercontent.com/gmtanner-cord/DATA470-2024/c39c51759761bff7abd5b42e5ab9ff35bf524515/fmhousing/FM_Housing_2018_2022_clean.csv"
)      
FM_housing <- FM_housing %>%
  rename(
    year_built = `Year Built`,
    total_sqft = `Total SqFt.`,
    sold_price = `Sold Price`
  )

set.seed(42) # This is for reproducibility, so that everyone gets the same answers.
test_index <- createDataPartition(FM_housing$`sold_price`, p = 0.20, list = FALSE)
test_set <- FM_housing[test_index,]
train_set <- FM_housing[-test_index,]
```

## Data Preprocessing

```{r echo = FALSE}
# Center and scale 
scaling_vars <- train_set %>%
  select(`year_built`, `total_sqft`) %>%
  preProcess(method = c("center", "scale"))

train_scaled <- predict(scaling_vars, select(train_set, `year_built`, `total_sqft`)) %>%
  as.data.frame() %>%
  cbind(train_set %>%
           select(`sold_price`))

# Center and scale in the test set
test_scaled <- predict(scaling_vars, select(test_set, `year_built`, `total_sqft`)) %>%
  as.data.frame() %>%
  cbind(test_set %>%
           select(`sold_price`))

```

## Train the kNN Model

```{r echo = FALSE}
# Define cross-validation control
control <- trainControl(method = "cv", number = 10)

# Train kNN model with cross-validation
model_knn_cv <- train(
  `sold_price` ~ .,
  data = train_scaled,
  method = "knn",
  tuneGrid = expand.grid(k = seq(1, 81, 4)), # Adjust `k` range as needed
  trControl = control
)

# Print the model
model_knn_cv

```

## Model Performance

```{r echo = FALSE}
# Evaluate model performance
y_hat_knn_cv <- predict(model_knn_cv, newdata = test_scaled)
postResample(y_hat_knn_cv, test_scaled$`sold_price`)

```

## Performance Metrics
```{r echo = FALSE}
# RMSE (Root Mean Squared Error)
rmse <- sqrt(mean((y_hat_knn_cv - test_scaled$sold_price)^2))

# R-squared
rsq <- cor(y_hat_knn_cv, test_scaled$sold_price)^2





```
#### RMSE
```{r echo = FALSE}
rmse
```
#### RSQ

```{r echo = FALSE}
rsq
```


# Visualizations

I generated two main plots to evaluate the performance of my model:

Residuals Distribution: This plot shows the distribution of the residuals (the difference between predicted and actual sold prices).
Actual vs Predicted Sold Price: This scatter plot compares the actual and predicted sold prices.

## Residuals Distribution

```{r echo = FALSE}
# Generate residuals plot
library(ggplot2)

residuals <- y_hat_knn_cv - test_scaled$sold_price
residual_plot <- ggplot(data = data.frame(residuals), aes(x = residuals)) +
  geom_histogram(binwidth = 50, fill = "skyblue", color = "black") +
  ggtitle("Residuals Distribution") +
  xlab("Residuals") +
  ylab("Frequency")

residual_plot

```

## Actual vs Predicted Sold Price

```{r echo = FALSE}
# Generate actual vs predicted plot
prediction_plot <- ggplot(data = data.frame(actual = test_scaled$sold_price, predicted = y_hat_knn_cv), aes(x = actual, y = predicted)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  ggtitle("Actual vs Predicted Sold Price") +
  xlab("Actual Sold Price") +
  ylab("Predicted Sold Price")

prediction_plot

```

